---
title: "Midterm Fall 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Orchid Species Richness along Elevational and Environmental Gradients in Yunnan, China (2015)

https://doi.org/10.1371/journal.pone.0142621

In this paper, Zhang et al. seek to understand environmental and geographical effects on Orchids in Yunan, China. The scientists used R and data from the Chinease Virtual Herbarium and aimed to understand the relationship between species richness and elevation. Further, they looked to examine how elevational diversity in Yunnan Province, China, might be explained by mid-domain effect (MDE), species–area relationship (SAR), water–energy dynamics (WED), Rapoport’s Rule, and climatic variables.For the simplicity of this data replication project, I will recreate the graphs that explain the species composition and orchid habitats which help explain the results more clearly and graphically(Figures 1 & 2). Secondly, I will recreate the graphs that explain elevational patterns of orchid richness in Yunan (Figure 3a & 3b). 

## Establishing libraries and loading the data set:

```{r} 
library(readr)
library(readxl)
library(tidytext)
library(tidyverse)
library(dplyr)



f <- "Midterm data.xlsx"
d <- read_excel(f)
head(d)
```

This loads the data that was appended to the article. The PDF supplied was then transfered into an .xlsx file by using the application Tabula. Because of the methodology of Tabula, there may be small misalignments of data going forward with this replication. 







## Species composition and orchid habitats in Yunnan (Part 1):

From the paper the following data was extracted to compare the results I will derive from my analysis:

Number of specimens collected = 4202
  Epiphytic - 47.12%
  Terrestial - 50.59%
  Saprophytic - 2.28%
             
              
Number of species = 691
  Epiphytic - 364 (52.68%)
  Terrestial - 305 (44.14%)
  Saprophytic - 22 (3.18%)


```{r} 

#figure1
d$`Life form`<- tolower(d$`Life form`)
Specimen_breakdown <- group_by(d, `Life form`) %>% summarize(count = n()) %>% arrange(desc(count)) %>%
  mutate(percentage = count/sum(count)*100)
Specimen_breakdown #gives statistical analysis of Life form by specimen


#run here
names <- Specimen_breakdown$`Life form`
counts <- Specimen_breakdown$count
barplot(counts, names.arg = names, main = "# of Specimens by Life form")

d$Species <- tolower(d$Species)
Species_breakdown <- select(d, Species, 'Life form') %>% unique() %>% 
  group_by(`Life form`) %>% summarize(count =n()) %>% arrange(desc(count)) %>%
  mutate(percentage = count/sum(count)*100) #give statistical analysis of Life form by species
Species_breakdown

#runhere
names <- Species_breakdown$`Life form`
counts <- Species_breakdown$count
barplot(counts, names.arg = names, main = "# of Species by Life form" )
```

As you can see, the numbers that were extracted from the paper are very similar to the ones that are explained by the breakdown for both Specimen and Species. This can be explained by the misalignment of data that was transferred initially. 

Now, to do the analysis of number of specimen and species by habitat type.

```{r}
#figure2
d$Habitat <- tolower(d$Habitat)
Habitat_specimen <- group_by(d, Habitat) %>% summarize(count = n()) %>% arrange(desc(count)) %>%
  mutate(percentage = count/sum(count)*100)
Habitat_specimen

names <- Habitat_specimen$Habitat
counts <- Habitat_specimen$count
barplot(counts, names.arg = names, las = 2) 

```

And for species ...

```{r}

d$Species <- tolower(d$Species)
Habitat_species <- select(d, Species, Habitat) %>% unique() %>%
group_by(Habitat) %>% summarize(count = n()) %>% arrange(desc(count)) %>%
  mutate(percentage = count/sum(count)*100)
Habitat_species

names <- Habitat_species$Habitat
counts <- Habitat_species$count
barplot(counts, names.arg = names, las = 2)

```



## Elevational patterns of orchid richness in Yunnan (Part 2a) 

For this analysis, I will have to create Bins to evaluate the number of species within each elevation bin. Like Zhang et al., I will designate at 200 m interval and do the analysis from there. 

```{r}

##figure3a
colnames(d)
v <- select(d, Species, elevation, 'Life form')
  summary(v)
v
head(v)

breaks <- c(0, 300, 500, 700, 900, 1100, 1300, 1500, 1700, 1900, 2100, 2300, 2500, 2700,
            2900, 3100, 3300, 3500, 3700, 3900, 4100, 4300, 4500)
tags <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22)

group_tags <- cut(v$elevation,
                  breaks = breaks,
                  include.lowest = TRUE,
                  right = FALSE,
                  labels = tags)

v<- cbind(v,group_tags)

v <- cbind(v, genus <- sapply(str_split(v$Species," "), `[`, 1))
names(v)<- c("Species","Elevation","Life Form","Elevation Bin","Genus")

head(v)

```

Here, the breaks designate the intervals in the elevation where the specimen resides. Then assigns the bin which the specimen was assigned and merged it with v, the filtered database. Genus was also desigated by splitting the string of the species name and only kept the first term.

Now to create the scatterplot graphs for Specimen (T), Species(S), and Genus(C)...

```{r}
fig3aT <- group_by(v, `Elevation Bin`) %>% summarize(count=n())
names(fig3aT) <- c("Bin", "Count")
fig3aS <- group_by(v, `Elevation Bin`,Species) %>% summarize() %>% group_by(`Elevation Bin`) %>% summarize(count=n())
names(fig3aS) <- c("Bin", "Count")
fig3aC <- group_by(v, `Elevation Bin`,Genus) %>% summarize() %>% group_by(`Elevation Bin`) %>% summarize(count=n())
names(fig3aC) <- c("Bin", "Count")

x <- fig3aT$Bin
yT <- fig3aT$Count
xT <- as.numeric(levels(x))[x]

x <- fig3aS$Bin
yS <- fig3aS$Count
xS <- as.numeric(levels(x))[x]

x <- fig3aC$Bin
yC <- fig3aC$Count
xC <- as.numeric(levels(x))[x]

fitT <- lm(yT ~ xT + I(xT^2) + I(xT^3))
fitS <- lm(yS ~ xS + I(xS^2) + I(xS^3))
fitC <- lm(yC ~ xC + I(xC^2) + I(xC^3))

plot(fig3aT, main = "Species Richness", )
points(xT, fitted(fitT), pch = 20, col = 'red')
quadraticT <- lines(sort(xT), fitted(fitT)[order(xT)], col='red', type='b')

points(xS, fitted(fitS), pch = 20, col = 'blue')
quadraticS <- lines(sort(xS), fitted(fitS)[order(xS)], col='blue', type='b')

points(xC, fitted(fitC), pch = 20, col = 'black')
quadraticC <- lines(sort(xC), fitted(fitC)[order(xC)], col='black', type='b')
```

Unfortunately, I was unable to plot the actual values for the Species and Genus plots, only the fitted polynomial regression. 

## Summary for Specimen
```{r}
summary(fitT)
```
In comparison, the R squared value for Zhang et al. is: 0.902

## Summary for Species
```{r}
summary(fitS) 
```
In comparison, the R squared value for Zhang et al. is: 0.964

## Summary for Genus
```{r}
summary(fitC) 
```
In comparison, the R squared value for Zhang et al. is: 0.951



## Elevational patterns of orchid richness in Yunnan (Part 2b) 

```{r}
#figure3b

fig3bT <- group_by(v, `Elevation Bin`,`Life Form`, Species) %>% summarize() %>% filter(`Life Form` == "epiphyte") %>%
  group_by(`Elevation Bin`) %>% summarize(count=n())
names(fig3bT) <- c("Bin", "Count")
fig3bS <- group_by(v, `Elevation Bin`,`Life Form`, Species) %>% summarize() %>% filter(`Life Form` == "terrestrial") %>%
  group_by(`Elevation Bin`) %>% summarize(count=n())
names(fig3bS) <- c("Bin", "Count")
fig3bC <- group_by(v, `Elevation Bin`,`Life Form`, Species) %>% summarize() %>% filter(`Life Form` == "saprophyte") %>%
  group_by(`Elevation Bin`) %>% summarize(count=n())
names(fig3bC) <- c("Bin", "Count")


x <- fig3bT$Bin
yT <- fig3bT$Count
xT <- as.numeric(levels(x))[x]

x <- fig3bS$Bin
yS <- fig3bS$Count
xS <- as.numeric(levels(x))[x]

x <- fig3bC$Bin
yC <- fig3bC$Count
xC <- as.numeric(levels(x))[x]

fitT <- lm(yT ~ xT + I(xT^2) + I(xT^3))
fitS <- lm(yS ~ xS + I(xS^2) + I(xS^3))
fitC <- lm(yC ~ xC + I(xC^2) + I(xC^3))


plot(fig3bT, main = "Species Richness - Life form")
quadraticT <- lines(sort(xT), fitted(fitT)[order(xT)], col='red', type='b')

points(xS, fitted(fitS), pch = 20, col = 'blue')
quadraticS <- lines(sort(xS), fitted(fitS)[order(xS)], col='blue', type='b')

points(xC, fitted(fitC), pch = 20, col = 'black')
quadraticC <- lines(sort(xC), fitted(fitC)[order(xC)], col='black', type='b')
```

This graph is very similar to the Figure 3b presented by Zhang et al. Here is the summary breakdown and comparison of R squared values.

```{r}
summary(fitT)
```
The R squared value for Epiphytic presented by Zhang was 0.954

```{r}
summary(fitS) 
```
The R squared value for Terrestrial presented by Zhang was 0.951

```{r}
summary(fitC)
```
The R squared value for Saprophytic presented by Zhang was 0.317






